### 2017-12-23 abstract.tex
failure change

%%%% Time-stamp: <2013-02-25 10:31:01 vk>

\section*{Abstract}
\label{cha:abstract}

This is a placeholder for the abstract. It summarizes the whole thesis
to give a very short overview. Usually, this the abstract is written
when the whole thesis text is finished.

%\glsresetall %% all glossary entries should be used in long form (again)
%% vim:foldmethod=expr
%% vim:fde=getline(v\:lnum)=~'^%%%%\ .\\+'?'>1'\:'='
%%% Local Variables:
%%% mode: latex
%%% mode: auto-fill
%%% mode: flyspell
%%% eval: (ispell-change-dictionary "en_US")
%%% TeX-master: "main"
%%% End: