### 2017-12-24 article.tex
Add other sections

%%%% Time-stamp: <2012-08-20 17:41:39 vk>

%% example text content
%% scrartcl and scrreprt starts with section, subsection, subsubsection, ...
%% scrbook starts with part (optional), chapter, section, ...

\section{Annual Report}
\subsection{Overview}
JD.com, Inc’s Annual Report for FY 2016 comprises the mainly three parts:
\begin{enumerate}
	\item[\uppercase\expandafter{\romannumeral1}] This part includes corporate information, introducing the development of the company, the basic strategy, and brief operating and financial situation during the last year. It also includes the information about its directors, managers and employees. And also the strategy towards them is included.
	\item[\uppercase\expandafter{\romannumeral2}] This part introduces the details of the company’s dividend policy. It also includes the adjustments to statements and its principal of accounting.
	\item[\uppercase\expandafter{\romannumeral3}] The third part is mainly about JD’s financial statements and exhibits of materials.
	\item[\uppercase\expandafter{\romannumeral4}] This part is JD’s financial statements and its principal to record.
\end{enumerate}

JD.com, Inc. annual report for FY 2016 provides a detailed and comprehensive review of the operating and financial performance of the issuer of the annual report and its principal subsidiaries in the last financial year.

\subsection{External users}
External users of the annual report include present and potential investors (equity owners), employees of the company, lenders of funds, customers, suppliers, government agencies and other members of the public: 
\begin{itemize}
\item Investors. The providers of risk capital and their advisers are concerned with the risk inherent in, and return provided by, their investments.  They need information to help them determine whether they should buy, hold or sell.  Shareholders are also interested in information that enables them to assess the ability of the enterprise to pay dividends. 
\item Employees. Employees and their representative groups are interested in information about the stability and profitability of their employers.  They are also interested in information which enables them to assess the ability of the enterprise to provide remuneration, retirement benefits and employment opportunities. 
\item Lenders. Lenders are interested in information that enables them to determine whether their loans, and the interest attaching to them, will be paid when due. 
\item Suppliers and other trade creditors. Suppliers and other creditors are interested in information that enables them to determine whether amounts owing to them will be paid when due. Trade creditors are likely to be interested in an enterprise over a shorter period than lenders unless they are dependent upon the continuation of the enterprise as a major customer. 
\item Customers. Customers have an interest in information about the continuance of an enterprise, especially when they have a long-term involvement with, or are dependent on, the enterprise. 
\item Governments and their agencies. Governments and their agencies are interested in the allocation of resources and, therefore, the activities of enterprises.  They also require information in order to regulate the activities of enterprises, determine taxation policies and as the basis for national income and similar statistics. 
\item Public. Enterprises affect members of the public in a variety of ways.  For example, enterprises may make a substantial contribution to the local economy in many ways including the number of people they employ and their patronage of local suppliers.  Financial statements may assist the public by providing information about the trends and recent developments in the prosperity of the enterprise and the range of its activities.
\end{itemize}
While the objective of annual reports is to provide information about the financial position, performance and changes in financial position of an enterprise, it cannot cater to all the information needs of users. However, as investors are providers of risk capital to the enterprise, financial information from the annual report that meet their needs will also meet the common needs of other users.

\section{Corporate Governance}
\subsection{Board of Directors}

\begin{table}[H]	
\begin{center}
\begin{tabular}{ |l|l| }
	\hline
	\multicolumn{2}{|c|}{\textbf{Board of Directors}} \\
	\hline
	Richard Qiangdong Liu	&		Chairman and Chief Executive Officer \\
	\rowcolor[gray]{.95}
	Martin Chiping Lau		& 		Director\\
	Ming Huang				&		Independent Director \\
	\rowcolor[gray]{.95}
	Louis T. Hsieh			&		Independent Director \\
	David Daokui Li			&		Independent Director \\
	\hline
\end{tabular}
\end{center}
\caption{Board of directors of JD.com, Inc}\label{table:1}
\end{table}

\subsubsection{Appointment of New Directors}
\subsubsection{Directors’ Remuneration}
In 2016, JD paid an aggregate of approximately RMB12.7 million (US\$1.8 million) in cash to their executive officers, and approximately US\$0.2 million in cash to their non-executive directors. They have not set aside or accrued any amount to provide pension, retirement or other similar benefits to their executive officers and directors\footnote{See JD.com, Inc. annual report 2016, Page 115}.
The details of remuneration are not mentioned in the report.

\subsubsection{Directors’ Appointments in Other Institutions}
Martin Chiping Lau is the only director who also holds positions in other institutions.
The situation is as follows:

\begin{table}[H]	
	\begin{center}
		\begin{tabular}{ |l|l|l| }
			\hline
			\rowcolor[gray]{.95}
			\textbf{DIRECTOR}	& \textbf{COMPANY} &\textbf{APPOINTMENT} \\
			\hline
			 	            	& Tencent Holdings Limited & President and Executive Director\\			 	            	
			Martin Chiping Lau	& Leju Holdings Limited & Director \\
			 			        & Kingsoft Corporation Limited &Non-Executive Director \\
			\hline
		\end{tabular}
	\end{center}
	\caption{Directors’ Appointments in Other Institutions}\label{table:1}
\end{table}

\subsubsection{Directors’ Interests in Shares and Debentures}
The following table sets forth information with respect to the beneficial ownership of our ordinary shares as of February 28, 2017 by\footnote{See JD.com, Inc. annual report 2016, Page 122}:
\begin{itemize}
	\item each of our directors and executive officers; and
	\item each person known to us to own beneficially more than 5\% of our total outstanding shares.
\end{itemize}

\begin{table}[H]	
	\begin{center}
		\begin{tabular}{|l|p{2.05cm}|c|p{2.05cm}|c|c}\hline
			        &\multicolumn{2}{c|}{DIRECT INTEREST}&\multicolumn{2}{c|}{DEEM INTEREST}\\
			        \hline
			Director&Beginning of Year&End of Year&Beginning of Year&End of Year\\
			\hline
			Richard Qiangdong Liu&452,044,989&452,044,989& &\\
			\hline
		\end{tabular}
	\end{center}
	\caption{Board of directors of JD.com, Inc}\label{table:1}
\end{table}

\subsubsection{Share Options}
A director is not required to hold any shares in JD’s company by way of qualification. A director who is in any way, whether directly or indirectly, interested in a contract or transaction or proposed contract or transaction with the company must declare the nature of his interest at a meeting of the directors.

Under JD’s current memorandum and articles of association, our board of directors will not be able to form a quorum without Mr. Richard Qiangdong Liu for so long as Mr. Liu remains a director.(Page118)

\subsubsection{Audit Committee}
JD’s audit committee consists of Louis T. Hsieh, Ming Huang and David Daokui Li. Mr. Hsieh is the chairman of their audit committee.
The audit committee oversees JD’s accounting and financial reporting processes and the audits of the financial statements of our company.
The audit committee is responsible for, among other things:
\begin{itemize}
\item appointing the independent auditors and pre-approving all auditing and non-auditing services permitted to be performed by the independent auditors;
\item reviewing with the independent auditors any audit problems or difficulties and management’s response;
\item discussing the annual audited financial statements with management and the independent auditors;
\item reviewing the adequacy and effectiveness of our accounting and internal control policies and procedures and any steps taken to monitor and control major financial risk exposures;
\item reviewing and approving all proposed related party transactions;
\item meeting separately and periodically with management and the independent auditors;
\item monitoring compliance with our code of business conduct and ethics, including reviewing the adequacy and effectiveness of our procedures to ensure proper compliance.
\end{itemize}

\subsubsection{Nominating and Corporate Governance Committee}
JD’s nominating and corporate governance committee consists of David Daokui Li and Louis T. Hsieh. Mr. Li is the chairperson of their nominating and corporate governance committee. The nominating and corporate governance committee is responsible for, among other things:
\begin{itemize}
\item	selecting and recommending to the board nominees for election by the shareholders or appointment by the board;
\item	reviewing annually with the board the current composition of the board with regards to characteristics such as independence, knowledge, skills, experience and diversity;
\item	making recommendations on the frequency and structure of board meetings and monitoring the functioning of the committees of the board; and
\item	advising the board periodically with regards to significant developments in the law and practice of corporate governance as well as our compliance with applicable laws and regulations, and making recommendations to the board on all matters of corporate governance and on any remedial action to be taken.
\end{itemize}

\subsection{Internal Control}
Internal control assists in the company’s effective and efficient operations by helping it to react suitably to significant business, operational, financial, compliance and other risks to achieve the company’s aims. JD did not employ any internal auditors.

\subsection{Independent Auditors}
Mr. Hsieh, Mr. Huang and Mr. Li in the company’s audit committee satisfy the “independence” requirements of NASDAQ and Rule 10A-3 under the Securities Exchange Act of 1934. It provides service as states in 3.1.

\section{Operating Activities}
The principal activities of the Company are online direct sales, online marketplace and other services. Its subsidiaries are principally set to assistant their operating activities, all of which were built in China.
The Company comprises the following branches and subsidiaries:
\begin{itemize}
\item Jingdong Century-Established in April 2007, and certain of its subsidiaries in China, which primarily engage in retail business;
\item	Shanghai Shengdayuan Information Technology Co., Ltd., or Shanghai Shengdayuan-Established in April 2011 and primarily operates their online marketplace business;
\item	Tianjin Star East Corporation Limited, or Star East-Established in April 2012 and provides primarily warehousing and related services;
\item	Beijing Jingbangda Trade Co., Ltd., or Jingbangda-Established in August 2012 and provides primarily courier services.
\end{itemize}

The success of this company hinges on their ability to provide superior customer experience. Their slogan is “selection, speed, quality, value”, and they are committed to optimizing customer experience and achieving customer satisfaction. They rely on online direct sale of computer, communication and consumer electronics for a significant portion of their net revenues.

The director of the Group admitted that they have incurred significant net losses and they may continue to experience significant losses in the future. It is also reported that if they are unable to provide superior customer experience, their business and reputation may be materially and adversely affected.

JD plan to further expand their fulfillment infrastructure. If they are not able to manage such expansion successfully, their growth potential, business and results of operations may be materially and adversely affected.

Cost of revenue account for the Group’s most cost and expense, and it decrease by 1.8\% during 2016. Fulfillment increases by 0.4\%, technology and content increases by 0.3\% and general and administrative increases by 0.2\%. While marketing decreases 0.2\% and impairment of goodwill and intangible assets decreases 1.5\%. This decrease in expense more than offset the increase in expense which resulted in loss decrease of 3.7\% as a percentage of revenue.

\section{Income Statement}

The Group’s income statement covers the period from 1 January 2014 to 31 December 2016.

Turnover represents online direct sales, services and others. The Group’s turnover is as follows:
\begin{table}[H]	
	\begin{center}
		\begin{tabular}{lcccc}
			\toprule
			 &\textbf{2015}&\textbf{2016}&\textbf{Absolute change}&\textbf{Ralative change}\\

			\midrule
	    	Turnover&181,275&260,122&78,847&43.5\%\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Turnover (in thousands of RMB \textyen)}\label{table:1}
\end{table}

The increase of 43.5\% in turnover from FY 2015 was achieved from an increase in online direct sales, which dominants about 38.6\%.

\subsection{Group Revenue According to Activities}

\begin{table}[H]	
	\begin{center}
		\begin{tabular}{lcc}
			\toprule
			\textbf{Fiscal year}&\textbf{2015}&\textbf{2016}\\
			\midrule
		    Online direct sales&167,721&237,702\\
			Services and others&13,554&22,420\\
			\qquad\emph{Total}&181,275&260,122\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Revenue (in thousands of RMB \textyen)}\label{table:1}
\end{table}

The Group’s online direct sales increases 41\% and services an others increase 65\%, resulting in 38.6\% increase in turnover. This illustrates a good state of operation during year 2016. We can also calculate that online direct sales generates the most profit for the company.

\subsection{Revenue Recognition}

·	The Group recognizes revenues from advertising placements ratably over the period during which the advertising services are provided or on the number of times that the advertisement has been displayed based on cost per thousand impressions, and recognizes revenues from pay for performance marketing services based on effective clicks.

·	The Group offers consumer financing to individual customers and supply chain financing to suppliers and merchants on the Group’s online marketplace. Revenues resulting from these financing services are recognized in accordance with the contractual terms.

·	The Group recognizes the revenues from the online direct sales on a gross basis as the Group is primarily obligated in these transactions, is subject to inventory risk, has latitude in establishing prices and selecting suppliers, or has met several but not all of these indicators.

·	The Group earns transaction fees from processing transactions for online payment customers. Services resulting from these transactions are recognized as service revenues when processing services are completed.

·	Advertising arrangements involving multiple deliverables are allocated into single-element arrangements based on their relative selling price in the absence of both vendor specific objective evidence and third party evidence, and the related revenue is recognized over the period during which the element is provided.

\subsection{Common Size Income Statement}
The Common Size Income Statement is as follows:

\section{Taxes}
\subsection{Taxation Summary}

The Group is imposed on the following three sorts of taxes:
\begin{itemize}
	\item \textbf{Value added tax.}
    During the periods presented, the Group is subject to statutory VAT\footnote{A pilot program for transition from business tax to value added tax (“VAT”) for certain services revenues was launched in Shanghai on January 1, 2012.} rate of 13\% and 17\% for revenues from sales of audio, video products and books and sales of other products, respectively, in the PRC. 
	\item \textbf{Business tax.}
	Before May 1, 2016, the Group was subject to 5\% business tax and related surcharges for revenues from online payment services. Starting from May 1, 2016, the Group is subject to 6\% VAT for revenues from online payment	services. Business tax and the related surcharges are recognized when the revenue is earned.
	\item \textbf{Income tax.}
	Three subsidiaries of the Group, Beijing Shangke, Chongqing Haijia and Chengdu Century, enjoyed a preferential
	income tax rate of 15\%. The Group’s other PRC subsidiaries, VIEs\footnote{A variable interest entity (VIE) is an entity that an investor has a controlling interest in, but this controlling interest is not based on a majority of voting rights.} and VIEs’ subsidiaries are subject to the statutory income tax rate of 25\%. 
\end{itemize}
As a  matter of fact, JD.com attempts to obtain tax allowance and exemption by being classified as “high and new technology enterprises”. 
\subsection{Income tax}
The amounts of income tax benefits or expenses for the year ended December 31,2015 and the year ended December 31,2016 are tabulated as follows\footnote{See JD.com, Inc. annual report 2016, Page 130}:

\begin{table}[H]	
	\begin{center}
		\begin{tabular}{lcc}
			\toprule
			\textbf{Fiscal year}&\textbf{2015}&\textbf{2016}\\
			\midrule
			Current income tax expenses&(28,322)&(214,282)\\
			Deferred tax benefits&\underline{42,584}&\underline{34,782}\\
			\qquad\emph{Income tax benefits/(expenses)}&\uuline{14,262}&\uuline{(179,500)}\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Income tax benefits or expenses (in thousands of RMB \textyen)}\label{table:1}
\end{table}

Current income taxes are provided on the basis of net income for financial reporting purposes, adjusted for income and expense items which are not assessable or deductible for income tax purposes, in accordance with the regulations of the relevant tax jurisdictions. The Group follows the liability method of accounting for income taxes. Under this method, deferred tax assets and liabilities are determined based on the temporary differences between the financial statements carrying amounts and tax bases of existing assets and liabilities by applying enacted statutory tax rates that will be in effect in the period in which the temporary differences are expected to reverse.

The Group records a valuation allowance to reduce the amount of deferred tax assets if based on the weight of available evidence, it is more-likely-than-not that some portion, or all, of the deferred tax assets will not be realized. The effect on deferred taxes of a change in tax rates is recognized in the Consolidated Statements of Operations and Comprehensive Loss in the period of change. The Group recognizes in its consolidated financial statements the benefit of a tax position if the tax position is “more likely than not” to prevail based
on the facts and technical merits of the position. Tax positions that meet the “more likely than not” recognition threshold are measured at the largest amount
of tax benefit that has a greater than fifty percent likelihood of being realized upon settlement. The Group estimates its liability for unrecognized tax benefits
which are periodically assessed and may be affected by changing interpretations of laws, rulings by tax authorities, changes and/or developments with
respect to tax audits, and expiration of the statute of limitations. The ultimate outcome for a particular tax position may not be determined with certainty
prior to the conclusion of a tax audit and, in some cases, appeal or litigation process. The actual benefits ultimately realized may differ from the Group’s
estimates. As each audit is concluded, adjustments, if any, are recorded in the Group’s consolidated financial statements in the period in which the audit is
concluded. Additionally, in future periods, changes in facts, circumstances and new information may require the Group to adjust the recognition and
measurement estimates with regard to individual tax positions. Changes in recognition and measurement estimates are recognized in the period in which the
changes occur. As of December 31, 2014, 2015 and 2016, the Group did not have any significant unrecognized uncertain tax positions.

\section{Dividends}
\subsection{Factors on Dividend Policy}
There are several factors that influence whether a company pays a dividend and how much it chooses to pay. Some of the most influential factors are listed as follows.
\begin{itemize}
	\item \textbf{Income stability.} 
	Income stability is one of the top factors in determining dividend policies. Specifically, established companies with stable, predictable income streams are more likely to pay dividends than companies with growing or volatile income.\\
	Newer and rapidly growing companies rarely pay dividends, as they prefer to invest their profits back into the company to fuel even more future growth. And, companies with unstable revenue streams often choose not to pay dividends, or pay small dividends in order to make sure the payout will be sustainable.\\
	It looks terrible to investors when companies are forced to suspend or reduce dividend payments, so most like to err on the side of caution when deciding to implement a new dividend, waiting for several years of stable profits before doing so.

	\item  \textbf{Profitability.}
	Another factor that can influence management's dividend policies is the potential for better returns through capital reinvestment. In other words, if a company feels that it would be in the best interest of its shareholders to use its profits for other business activities besides paying dividends, it could choose not to pay – even if its revenues are stable and predictable.\\
	One great example of this is Warren Buffett's Berkshire Hathaway, which has never paid a dividend. Instead, Buffett feels that reinvesting the company's profits is a far better idea -- and he's been right. Berkshire has produced phenomenal returns for decades, and a big reason was the compounding effect of reinvesting its profits instead of paying them out.

	\item \textbf{Taxes.}
	Dividends are effectively taxed twice -- once at the corporate level, and again when they are paid out to shareholders.\\
	Because of this, many companies (and their investors) feel that other methods of returning capital, such as share repurchases, are a better way to go. Repurchasing shares has the same net effect as a dividend payment -- the intrinsic value of the company's shares increases as the share count drops. However, this can allow investors who like to reinvest their dividends to do so without having to worry about dividend taxes.

	\item \textbf{Legal requirements.} 
	It's also worth noting that some companies have no choice but to pay dividends. For example, real estate investment trusts receive some pretty nice tax benefits, but they are legally required to pay at least 90\% of their income to shareholders.\\
	On the other hand, some companies need to obtain approval before paying or increasing their dividends. Since the financial crisis, many banks need to submit capital plans for regulatory approval for any plans to boost their payouts.

	\item \textbf{Economic conditions.}
	Finally, another major factor that influences dividend policies is the market environment. If a certain sector is having trouble and anticipates profits falling, it's common for companies to get quite defensive when it comes to their dividends.\\	
	This can be seen currently in the energy sector, where low oil prices have wreaked havoc on many companies' profitability, which has led to several major companies slashing their dividends recently.
\end{itemize}

As for JD.com, Inc., it stated that the form, frequency and amount of dividend payment will depend upon its future operations and earnings, capital requirements and surplus, general financial condition, contractual restrictions and other factors that the board of directors may deem\footnote{See JD.com, Inc. annual report 2016, Page 130}.

\subsection{Dividend Policy performed by JD.com}
JD.com, Inc. does not reflect an active attitudes towards dividend payment. So far JD.com have not declared or paid any dividends on ordinary shares, nor do JD.com have any present plan to pay any cash dividends on ordinary shares in the foreseeable future\footnote{See JD.com, Inc. annual report 2016, Page 131}.

\section{Share Capital and Reserves}

JD.com, Inc. holds outstanding shares consisting of ordinary shares, series A and A-1 convertible preferred shares and series B convertible preferred shares\footnote{See JD.com, Inc. annual report 2016, Page F-10}.

\begin{table}[H]	
	\begin{center}
		\begin{tabular}{lrr}
			\toprule
			\textbf{Fiscal year}&\textbf{2015}&\textbf{2016}\\
			\midrule
			Cash and cash equivalents&17,863,868&19,771,695\\
			Restricted cash&2,114,913&4,391,955\\
			Short-term investments&2,780,482&4,391,955\\
			Investment securities&-&723,449\\
			Accounts receivable, net&8,193,665&17,464,408\\
			Advance to suppliers&927,177&1,423,736\\
			Inventories, net&20,539,543&28,909,438\\
			Loan receivables, net&3,698,488&12,697,915\\
			Other investments&-&10,766,920\\
			Prepayments and other current assets&1,486,441&2,198,906\\
			Amount due from related parties&\underline{863,516}&\underline{1,410,050}\\
			\qquad\emph{Total current assets}&\uuline{58,468,093}&\uuline{106,932,098}\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Current assets (in thousands of RMB \textyen)}\label{table:1}
\end{table}

\section{Current Assets and Current Liabilities}
\subsection{Current Assets}
A current asset is any asset which can reasonably be expected to be sold, consumed, or exhausted through the normal operations of a business within the current fiscal year. JD.com records 11 current assets accounts in its consolidated balance Sheet\footnote{See JD.com, Inc. annual report 2016, Page F-3}.

We extract the following breakdown from JD.com, Inc. consolidated balance Sheet as of December 31, 2015 and as of December 31, 2016.\\

\begin{table}[H]	
\begin{center}
	\begin{tabular}{lrr}
		\toprule
		\textbf{Fiscal year}&\textbf{2015}&\textbf{2016}\\
		\midrule
		Cash and cash equivalents&17,863,868&19,771,695\\
		Restricted cash&2,114,913&4,391,955\\
		Short-term investments&2,780,482&4,391,955\\
		Investment securities&-&723,449\\
		Accounts receivable, net&8,193,665&17,464,408\\
		Advance to suppliers&927,177&1,423,736\\
		Inventories, net&20,539,543&28,909,438\\
		Loan receivables, net&3,698,488&12,697,915\\
		Other investments&-&10,766,920\\
		Prepayments and other current assets&1,486,441&2,198,906\\
		Amount due from related parties&\underline{863,516}&\underline{1,410,050}\\
		\qquad\emph{Total current assets}&\uuline{58,468,093}&\uuline{106,932,098}\\
		\bottomrule
	\end{tabular}
\end{center}
	\caption{Current assets (in thousands of RMB \textyen)}\label{table:1}
\end{table}
As we can see, the current assets jumped from \textyen58,468,093,000 to \textyen106,932,098,000.

\subsection{Current Liabilities}

We extract the following breakdown from JD.com, Inc. consolidated balance Sheet.\\

\begin{table}[H]	
	\begin{center}
		\begin{tabular}{lrr}
			\toprule
			\textbf{Fiscal year}&\textbf{2015}&\textbf{2016}\\
			\midrule
			Short-term borrowings&3,040,209&8,333,317\\
			Nonrecourse securitization debt&579,843&9,389,213\\
			Accounts payable&29,819,341&4,391,955\\
			Advance from customers&7,173,885&43,988,087\\
			Deferred revenues&983,721&11,632,766\\
			Taxes payable&103,211&1,221,865\\
			Amount due to related parties&104,726&575,848\\
			Accrued expenses and other current liabilities&\underline{7,178,065}&\underline{167,655}\\
			\qquad\emph{Total current liabilities}&\uuline{48,983,001}&\uuline{104,740,235}\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Current liabilities (In thousands of RMB \textyen)}\label{table:1}
\end{table}

It indicates that the current liabilities jumped from \textyen104,726,000 to \textyen104,740,235,000.

\subsection{Working Capital}
\begin{table}[H]	
	\begin{center}
		\begin{tabular}{lrr}
			\toprule
			\textbf{Fiscal year}&\textbf{2015}&\textbf{2016}\\
			\midrule
			Working capital&9,485,092&2,191,863\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Working capital (In thousands of RMB \textyen)}\label{table:1}
\end{table}

\subsection{Current Ratio}
\begin{table}[H]	
	\begin{center}
		\begin{tabular}{lrr}
			\toprule
			\textbf{Fiscal year}&\textbf{2015}&\textbf{2016}\\
			\midrule
			Current ratio&1.194&1.021\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Current ratio}\label{table:1}
\end{table}

\subsection{Quick Ratio}
\begin{table}[H]	
	\begin{center}
		\begin{tabular}{lrr}
			\toprule
			\textbf{Fiscal year}&\textbf{2015}&\textbf{2016}\\
			\midrule
			Quick ratio&29,819,341&4,391,955\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Quick ratio}\label{table:1}
\end{table}

\section{Inventories}
\subsection{Inventory Values}

Inventories, net, consist of the following\footnote{See JD.com, Inc. annual report 2016, Page F-48}:
\begin{table}[H]	
	\begin{center}
		\begin{tabular}{lrr}
			\toprule
			\textbf{Fiscal year}&\textbf{2015}&\textbf{2016}\\
			\midrule
			Products&20,436,275&28,778,528\\
			Packing materials and others&\underline{103,268}&\underline{130,910}\\
			\qquad\emph{Inventories, net}&\uuline{20,539,543}&\uuline{28,909,438}\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Inventories, net (in thousands of RMB \textyen)}\label{table:1}
\end{table}

\subsection{Inventory Valuation Policy}
Inventories, consisting of products available for sale, are stated at the lower of cost or market value\footnote{See JD.com, Inc. annual report 2016, Page F-20}. Cost of inventory is determined using the weighted
average cost method. Adjustments are recorded to write down the cost of inventory to the estimated market value due to slow-moving merchandise and
damaged goods, which is dependent upon factors such as historical and forecasted consumer demand, and promotional environment. The Group takes
ownership, risks and rewards of the products purchased, but has arrangements to return unsold goods with certain vendors. Write downs are recorded in cost of
revenues in the Consolidated Statements of Operations and Comprehensive Loss.

The Group also provides fulfillment-related services in connection with the Group’s online marketplace. Third-party sellers maintain ownership of their
inventories and therefore these products are not included in the Group’s inventories.
\subsection{Gross Profit Percentage on Sale}

Based on the data of JD.com, Inc. consolidated statements of operations and comprehensive loss\footnote{See JD.com, Inc. annual report 2016, Page F-48}, gross profit percentage on sale can be calculated as follows:

\begin{table}[H]	
	\begin{center}
		\begin{tabular}{lrr}
			\toprule
			\textbf{Fiscal year}&\textbf{2015}&\textbf{2016}\\
			\midrule
			Revenue&181,275,425&260,121,645\\
			Cost of goods sold&157,008,329&220,698,727\\
			Gross profit&24,267,096&39,422,918\\
			Gross profit percentage&13.39\%&15.16\%\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Gross profit percentage on sale (in thousands of RMB \textyen\ except for gross profit percentage)}\label{table:1}
\end{table}

\subsection{Inventory Turnover}
Based on the data of JD.com, Inc. consolidated balance sheet\footnote{See JD.com, Inc. annual report 2016, Page F-3 and JD.com, Inc. annual report 2015, Page F-3} and consolidated statements of operations and comprehensive loss\footnote{See JD.com, Inc. annual report 2016, Page F-48}, gross profit percentage on sale can be calculated as follows:

\begin{table}[H]	
	\begin{center}
		\begin{tabular}{lrr}
			\toprule
			\textbf{Fiscal year}&\textbf{2015}&\textbf{2016}\\
			\midrule
			Beginning inventory &12,190,843&20,539,543\\
			Ending inventory  &20,539,543&28,909,438\\
			Average inventory &16,365,193&24,724,490.5\\
			Cost of goods sold &157,008,329&220,698,727\\
			Inventory turnover &9.594&8.826\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Inventory turnover (in thousands of RMB \textyen\ except for inventory turnover)}\label{table:1}
\end{table}

\section{Accounts Receivable}
\subsection{Accounts receivable}

\begin{table}[H]	
	\begin{center}
		\begin{tabular}{lrr}
			\toprule
			\textbf{Fiscal year}&\textbf{2015}&\textbf{2016}\\
			\midrule
			Accounts receivable &8,465,644&17,880,720\\
			Allowance for doubtful accounts  &\underline{(271,979)}&\underline{(416,312)}\\
			\qquad\emph{Accounts receivable, net} &  \uuline{8,193,665}&\uuline{17,464,408)}\\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Accounts receivable (in thousands of RMB \textyen}\label{table:1}
\end{table}

\subsection{Avoidance and Reduction of Credit Losses} 
Accounts receivable are typically unsecured and are derived from revenues earned from customers in the PRC. The risk with respect to accounts receivable is mitigated by credit evaluations the Group performs on its customers and its ongoing monitoring process of outstanding balances.

In addition, JD Finance has changed from supporting the overall JD platform to an independently operated and self-funded business, and accounts receivables from consumer financing provided by JD Finance to consumers in marketplace are mainly for investment purpose and are reclassified to loan receivables since 2016.

In General, Companies are able to avoid and reduce credit losses through the following: 
Monitor and assess a customer’s creditworthiness through trade and bank references, and credit agencies (e.g. Credit Bureau Singapore), and only extend credit to those customers with good credit histories.

Companies may set a credit limit on each customer’s account which must not be exceeded.

If customers are not paying, second and third statements should be sent out to remind customers.

Companies could also opt to purchase business credit insurance to substantially reduce the risk of exposure to non-payment, and the accompanying bad debt loss. Commercial credit risk coverage can be written to include all customers, or it may be targeted to cover only certain buyers.

\section{Long-Term Assets, Impairment and Depreciation} 
\subsection{Property, equipment and software}

\begin{table}[H]	
	\begin{center}
		\begin{tabular}{lrrr}
			\toprule
			&\textbf{2015}&\textbf{2016}&\textbf{Change}\\
			\midrule
			Electronic equipment &	3,089,415&	4,422,833&	1,333,418\\
			Office equipment  &	202,422&	266,411&	63,989\\
			Vehicles&	843,922&	1,091,556&	247,634\\
			Logistic and warehouse equipment &	1,105,245&	1,578,653&	473,408\\
			Leasehold improvement&	306,352&	496,887&	190,535\\
			Software& 	149,590&	196,713&	47,123\\
			Building and building improvement & 	2,526,948&	3,069,651&	542,703\\
			Total  &	8,223,894&	11,122,704&	2,898,810\\
			Less: Accumulated depreciation&	(1,990,788)&	(3,725,675)	& \\
			\qquad\emph{Net book value} &	6,233,106&	7,397,029& \\
			\bottomrule
		\end{tabular}
	\end{center}
	\caption{Property, equipment and software receivable (in thousands of RMB \textyen)}\label{table:1}
\end{table}

The amount of Property, equipment and software has increased from FY 2015 to FY 2016 by RMB 2,898,810. It’s because all of the items are increasing, especially Electronic equipment.

\newpage

\section*{Appendix}
\addcontentsline{toc}{section}{Appendix}
yu

%\section{Example Chapter}
%
%This is my text with an example Figure~\ref{fig:example} and example
%citation~\cite{StrunkWhite} or \textcite{Bringhurst1993}. And there is another
%\enquote{citation} which is located at the bottom\footcite{tagstore}.
%
%\myfig{TU_Graz_Logo}%% filename in figures folder
%  {width=0.1\textwidth,height=0.1\textheight}%% maximum width/height, aspect ratio will be kept
%  {Example figure.}%% caption
%  {}%% optional (short) caption for table of figures
%  {fig:example}%% label
%
%Now you are able to write your own document. Always keep in mind: it's
%the \emph{content} that matters, not the form. But good typography is
%able to deliver the content much better than information set with bad
%typography. This template allows you to focus on writing good content
%while the form is done by the template definitions.

%% vim:foldmethod=expr
%% vim:fde=getline(v\:lnum)=~'^%%%%\ .\\+'?'>1'\:'='
%%% Local Variables: 
%%% mode: latex
%%% mode: auto-fill
%%% mode: flyspell
%%% eval: (ispell-change-dictionary "en_US")
%%% TeX-master: "main"
%%% End: